\\Large Rising temperatures alter how and where boreal mammals move

### 2025-11-11 writing/manuscript.Rmd
rm KS as author; rm extra from main text
Kirk asked to be removed because he didn't believe his contributions were sufficient for authorship

---
title: "\\Large Rising temperatures alter how and where boreal mammals move"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [braes, biol]
  - name: Christen H. Fleming
    institute: [ucf, smith]
  - name: Karen E. Hodges
    institute: [braes, biol]
  - name: Siobhan Darlington
    institute: [braes, biol]
  - name: Adam T. Ford
    institute: [braes, biol]
  - name: TJ Gooliaff
    institute: [forest]
  - name: Kirk Safford
    institute: bcp
  - name: Robert Serrouya
    institute: [braes, biol, biodiv]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - forest: British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development, Penticton, BC, Canada.
  - bcp: British Columbia Ministry of Environment and Parks, Penticton, BC, Canada.
  - biodiv: Wildlife Science Centre, Biodiversity Pathways, University of British Columbia Okanagan, Revelstoke, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'bc-mammals-temperature.bib'
csl: 'global-change-biology.csl'
# csl: 'movement-ecology.csl' # to count refs
fontsize: 12pt
indent: true
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=footnotesize, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage[left]{lineno} # for line numbers
    - \usepackage{booktabs} # to prevent tables from splitting
subparagraph: true # needed for \usepackage[small]{titlesec}
urlcolor: cyan
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for Delta symbol in table
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: true
editor_options: 
  chunk_output_type: console
---

@@

\noindent \textbf{Words in main text}: `r suppressMessages(wordcountaddin::word_count())`

@@

\noindent \textbf{References}: 165 (updated on 2025-10-31) <!-- counted using movement ecology csl -->

@@

@@

\noindent SM performed the data cleaning, ran the analyses, and wrote the manuscript. CHF wrote code for estimating instantaneous movement speeds. MJN conceived of the project idea and supervised SM throughout the project. All other authors contributed telemetry data and/or reviewed the interpretation of the results for their species of interest. KH and SD contributed substantially to manuscript review. All authors contributed to reviewing the manuscript.

>>>>

---
title: "\\Large Rising temperatures alter how and where boreal mammals move"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [braes, biol]
  - name: Christen H. Fleming
    institute: [ucf, smith]
  - name: Karen E. Hodges
    institute: [braes, biol]
  - name: Siobhan Darlington
    institute: [braes, biol]
  - name: Adam T. Ford
    institute: [braes, biol]
  - name: TJ Gooliaff
    institute: [forest]
  - name: Robert Serrouya
    institute: [braes, biol, biodiv]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - forest: British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development, Penticton, BC, Canada.
  - biodiv: Wildlife Science Centre, Biodiversity Pathways, University of British Columbia Okanagan, Revelstoke, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'bc-mammals-temperature.bib'
csl: 'global-change-biology.csl'
# csl: 'movement-ecology.csl' # to count refs
fontsize: 12pt
indent: true
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=footnotesize, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage[left]{lineno} # for line numbers
    - \usepackage{booktabs} # to prevent tables from splitting
subparagraph: true # needed for \usepackage[small]{titlesec}
urlcolor: cyan
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for Delta symbol in table
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: true
editor_options: 
  chunk_output_type: console
---

@@

\noindent \textbf{Words in main text}: `r suppressMessages(wordcountaddin::word_count()) - 80` <!-- does not include abstract or figure captions; "- 80" removes the stuff on page 2 -->

@@

\noindent \textbf{References}: 178 (updated on 2025-11-11, including +1 for @serrouya_ref) <!-- counted using movement ecology csl -->

@@

# Acknowledgements {-}

```{r}
ackn <- "\\noindent We would like to thank Kirk Safford from the British Columbia Ministry of Environment and Parks (BC Parks) in Penticton, BC, Canada for providing the mountain goat telemetry data. SM would also like to thank BC Parks and Biodiversity Pathways for financial suppport, and also all those who provided feedback following presentations and poster sessions."
```

`r ackn`

@@

```{r}
contrib <- "\\noindent SM performed the data cleaning, ran the analyses, and wrote the manuscript. CHF wrote code for estimating instantaneous movement speeds. MJN conceived of the project idea and supervised SM throughout the project. All other authors contributed telemetry data and/or reviewed the interpretation of the results for their species of interest. KH and SD contributed substantially to manuscript review. All authors contributed to reviewing the manuscript."
```

`r contrib`
